Sensitivity Analysis

### 2024-05-31 Sections/Results_Sensitivity.tex
31.05

	This work investigates the influence of pressure change on the state space and the extraction yield. The process model and parameters have been discussed in {\color{red}article 1}. The process model was calibrated on the set of experiments obtained at different operating conditions, $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. The sensitivity analysis has been performed assuming that the system operates at $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

>>>>

	%This work investigates the influence of pressure change on the state space and the extraction yield. The process model and parameters have been discussed in {\color{red}article 1}. The process model was calibrated on the set of experiments obtained at different operating conditions, $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. The sensitivity analysis has been performed assuming that the system operates in the middle of validate range, hence at $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

	This study examines the impact of pressure change on the state space and extraction yield. The process model and parameters are described in {\color{red}article 1}. Calibration of the process model was based on experiments performed under varying operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow between $3.33-6.67 \cdot 10^{-5}$ kg/s. A sensitivity analysis was conducted, assuming the system operates at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

### 2024-05-31 Sections/Sensitivity_Pressure.tex
31.05

	As discussed in Chapter \ref{CH:Governing_equations_chapter}, a small pressure wave propagates at the speed of sound relative to the flow. If the flow velocity is relatively low, all pressure changes are hydrodynamic (resulting from velocity motion) rather than thermodynamic. The Low Mach-number assumption leads to instant propagation of the thermodynamic pressure throughout the system. A single pressure value can be considered for the entire system, as all changes occur simultaneously throughout the device. A step-function shown in Figure \ref{fig:Sensitivty_P_P} illustrates the pressure change in the extractor.

@@

	The model is built such that the pressure change first affects $\frac{dP}{dt}$, and in the next time step, the energy balance through $\frac{\partial (P(t) A_f)}{\partial t}$, leading to the step change along the whole system as presented in Figure \ref{fig:Sensitivty_P_H}. The energy balance response becomes noticeable, as well as the change in the physical properties of the solvent. The uniform response across the entire extraction column length, represented by the homogeneously dark red colour, indicates that the whole system experiences an immediate and uniform change in enthalpy density in response to pressure changes.

	The pressure change affects the fluid's temperature inside the computational domain, but boundary values are subject to constraints specified at the extremes of that domain. Applying the Dirichlet boundary conditions assumes a fixed temperature value at the inlet, which can lead to a thermal gradient propagating along the system. In contrast, Neumann boundary conditions would dictate the heat flux at the boundaries. In this work, the Neumann boundary conditions equal to zero were applied to ensure that the temperatures at the inlet, the outlet, and the middle of the extractor are the same all the time.

@@

	Figure \ref{fig:Sensitivty_P_CS} shows the sensitivity of the solute concentration in the solid phase with respect to pressure change along a fixed bed in a supercritical fluid extraction process. As discussed in Chapter \ref{CH: Continuity}, the velocity of a fluid is inversely proportional to its density, which suggests that the pressure increase decreases the velocity. This leads to an extended residence time, ergo, a longer interaction between the solute and the solvent. Initially, the extraction process is in the kinetically-controlled regime, where the concentration gradient is high, and the limiting factor is the solute solubility. As discussed in ({\color{red}article 1}), the system is considered to be far from saturation, which can explain the low system response at the beginning of the process. The system response becomes more evident when the concentration gradient starts diminishing, and the extraction switches from the kinetic-controlled to the diffusion-controlled regime. The negative sign can be interpreted as a faster solute loss from the solid phase, which corresponds to enhanced mass transfer. Over time, the amount of solute becomes a limiting factor, and the pressure change has a lower effect on the system. Eventually, the sensitivities approach zero asymptotically when the solute is exhausted from the solid bed.

@@

	Figure \ref{fig:Sensitivty_P_CF} shows how sensitive is the solute concentration in the fluid phase to the pressure change. Although the pressure increase enhances the mass transfer, the system response is initially low, which reflects the idle period discussed above. Later, the sensitivities increase due to faster solute loss from the solid phase, which indicates that more solute is transported to the fluid phase. When the amount of solute in the solid phase becomes a limiting factor, the extraction rate slows down, sensitivities decline asymptotic towards zero, and the impact of the pressure change decreases. The advection causes the sensitivities to move along the system analogously to the solute in the fluid phase.

>>>>

	As discussed in Chapter \ref{CH:Governing_equations_chapter}, a small pressure wave propagates at the speed of sound relative to the flow. If the flow velocity is relatively low, all pressure changes are hydrodynamic (resulting from velocity motion) rather than thermodynamic. The Low Mach number assumption leads to instant propagation of the thermodynamic pressure throughout the system. A single pressure value can be considered for the entire system, as all changes occur simultaneously throughout the device. A step-function shown in Figure \ref{fig:Sensitivty_P_P} illustrates the pressure change in the extractor.

@@

	The model is built such that the pressure change first affects $\frac{dP}{dt}$, then in next time step, the energy balance is influenced through $\frac{\partial (P A_f)}{\partial t}$. This leads to the step change along the whole system as presented in Figure \ref{fig:Sensitivty_P_H}. The uniform response across the entire extraction column length, represented by the homogeneously dark red colour, indicates that the whole system experiences an immediate and uniform change in enthalpy density in response to pressure changes.

	The pressure change affects the fluid's temperature inside the computational domain, but boundary values are subject to constraints specified at the extremes of that domain. Applying the Dirichlet boundary conditions assumes a fixed temperature value at the inlet, which can lead to a thermal gradient propagating along the system. In contrast, Neumann boundary conditions would dictate the heat flux at the boundaries. In this work, the Neumann boundary conditions equal to zero were applied to ensure that the temperatures at the inlet, the outlet, and the middle of the extractor depends only on the pressure change.

@@

	Figure \ref{fig:Sensitivty_P_CS} shows the sensitivity of the solute concentration in the solid phase with respect to pressure change along a fixed bed in a supercritical fluid extraction process. As discussed in Chapter \ref{CH: Continuity}, the velocity of a fluid is inversely proportional to its density, which suggests that the pressure reduces the fluid's velocity. As a result, the residence time was extended, causing longer interaction between the solute and the solvent. Initially, the extraction process is in the kinetic-controlled regime, where the concentration gradient is high, and the limiting factor is the solute solubility. As discussed in ({\color{red}article 1}), the system is considered to be far from saturation, which can explain the low system response at the beginning of the process. The system response becomes more evident when the concentration gradient starts diminishing, and the extraction switches from the kinetic-controlled to the diffusion-controlled regime. The negative sign can be interpreted as a faster solute loss from the solid phase, which corresponds to enhanced mass transfer. Over time, the amount of solute becomes a limiting factor, and the pressure change effect on the system is reduced. Eventually, the sensitivities approach zero asymptotically when the solute is exhausted from the solid bed.

@@

	Figure \ref{fig:Sensitivty_P_CF} shows how sensitive is the solute concentration in the fluid phase to the pressure change. Compare to Figure \ref{fig:Sensitivty_P_CS}, the dynamic behaviour of the fluid phase sensitivities can be observed. The sensitivities moves along the system analogously to the solute in the fluid phase due to the advection.	Although the pressure increase enhances the mass transfer, the system response is initially low, which reflects the idle period discussed above. Later, the sensitivities increase due to faster solute loss from the solid phase, which indicates that more solute is transported to the fluid phase. When the amount of solute in the solid phase becomes a limiting factor, the extraction rate slows down and sensitivities decline asymptotic towards zero.
